---
title: "first_lake_report"
output:
  pdf_document: default
  html_document: default
date: "2025-03-18"
---

```{r setup, include=FALSE}
if (!require(dplyr)) {install.packages("dplyr")}
if (!require(tidyverse)) {install.packages("tidyverse")}

library(readr)
library(stringr)
```

## Downloading Folder with All Data\
```{r}
folder = "C:/Users/mvhfg/OneDrive - University of Missouri/North_Limnology/priv_lake_rep/Private_Lake_Reports/2024 Data"
files = list.files(
  path = folder,
  pattern = ".*csv$",
  ignore.case = F,
  full.names = T #provides path to folder
)

files2 <- list.files("2024 Data", pattern="*.csv")

library(readr)
library(stringr)

for(i in 1:length(files)) {
  file_name <- str_sub(string = files[i], start = 111)
  file_df <- read.csv(paste0(folder,"/",files2[i]),header = TRUE, sep=",")
  #file_df <- file_df[c(3:2714),2:3] #i only need the col 2&3 and keep rows 2:2714
  #cols <- names(file_df)[14] #getting the 14th column 
  #file_df[cols] <- lapply(file_df[cols], as.numeric) #making the cols column(s) numeric
  assign( x = file_name, value = file_df, envir = .GlobalEnv)
}
```


## Results

### Total Nitrogen & Total Phosphorus\
Total Phosphorus - Showing TP Data
```{r}
#all TP data
#we want to extract TP values for lake 14 - Tishomingo
# we are only looking at surface samples so the Depth has to be SURF
tp_14 <- SLAP_2024_TP.csv%>%
  filter(MULakeNumber == "014" & Depth == "SURF")
#for depth instead of looking for SURF, look for endDepth = 0
```

Averaging TP Data
```{r}
tp_14_avg <- tp_14 %>%
  group_by(Date)%>%
  summarise(Average_TP = mean(parameterValue)) #units in ug/L
```

Plot of TP
```{r}
tp_14_avg %>%
  ggplot( aes(x=Date, y=Average_TP)) +
  geom_hline(yintercept = 60.12, size = 2, color = "orchid4")+
  geom_line(aes(group=1), color = "gray76", size = 1.5)+
  geom_point(shape=21, color="black", fill="orchid4", size=8) +
    theme_bw()+
    theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())+
    xlab('Date')+
    ylab('TP Concentration (µg/L)')+
    labs(title = "Average Total Phosphorus Concentrations for Summer 2024")
```


Average TN Data
```{r}
tn_14 <- SLAP_2024_TN.csv%>%
  filter(MULakeNumber == "014" & endDepth == 0)

tn_14_avg <- tn_14 %>%
  group_by(Date)%>%
  summarise(Average_TN = mean(parameterValue)) #units in ug/L
```

Plot of TN
```{r}
tn_14_avg %>%
  ggplot( aes(x=Date, y=Average_TN)) +
  geom_hline(yintercept = 950.6, size = 2, color = "royalblue4")+
  geom_line(aes(group=1), color = "gray76", size = 1.5)+
  geom_point(shape=21, color="black", fill="royalblue4", size=8) +
    theme_bw()+
    theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())+
    xlab('Date')+
    ylab('TN Concentration (µg/L)')+
    labs(title = "Average Total Nitrogen Concentrations for Summer 2024")

#save in 600 by 400 png
```

### Toxins
```{r}
ANAtox <- SLAP_2024_ANA_tox.csv[,1:19] #keeps the first 19 columns
CLYtox <- SLAP_2024_CLY_tox.csv[,1:19]
MICtox <- SLAP_2024_MIC_tox.csv[,1:19]
SAXtox <- SLAP_2024_SAX_tox.csv[,1:19]

ANAtox_14 <- ANAtox%>%
  filter(MULakeNumber == "014")

CLYtox_14 <- CLYtox%>%
  filter(MULakeNumber == "014")

MICtox_14 <- MICtox%>%
  filter(MULakeNumber == "014")

SAXtox_14 <- SAXtox%>%
  filter(MULakeNumber == "014")

ALLtox_14 <- bind_rows(ANAtox_14, CLYtox_14, MICtox_14, SAXtox_14)  #units in ug/L
```

Full df of Toxins
```{r}
tox_avg_14 <- ALLtox_14 %>%
  group_by(Date, parameterType)%>%
  summarise(avg_tox = mean(parameterValue)) #units in ug/L

if (!require(ggplot2)) {install.packages("ggplot2")}


ggplot(tox_avg_14, aes(x=Date, y=avg_tox, group=parameterType)) +
  geom_line(size =0.75, aes(color =parameterType)) +
  geom_point(aes(shape=parameterType, color=parameterType, size=parameterType))+
  scale_shape_manual(values=c(15, 16, 17, 18))+
  scale_color_manual(values=c('brown4','orange', 'palegreen3', 'steelblue'))+
  scale_size_manual(values=c(8,8,5,6))+
  theme(legend.position="right")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())+
  xlab('Date')+
  ylab('Average Tox Concentration (µg/L)')+
  labs(title = "Cyanotoxin Concentrations for Summer 2024") 
#best one so far
```

### Secchi & CHL A

Secchi Depth Averages
```{r}
sec_14 <- SLAP_2024_SECCHI.csv%>%
  filter(MULakeNumber == "014")

sec_14_new <- sec_14 %>%
    mutate(endDepth = if_else(is.na(endDepth), 0, endDepth))

sec_14_avg <- sec_14_new %>%
  group_by(Date)%>%
  summarise(Average_SEC = mean(parameterValue))  #units in m
```

Plot of Secchi Depth
```{r}
sec_14_avg %>%
  ggplot( aes(x=Date, y=Average_SEC))+
  geom_hline(yintercept = 1.11, size = 2, color = "darkorange4")+
  geom_line(aes(group=1), color = "gray76", size = 1.5)+
  geom_point(shape=21, color="black", fill="darkorange4", size=8) +
    theme_bw()+
    theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())+
    xlab('Date')+
    ylab('Depth (m)')+
    labs(title = "Average Secchi Depths for Summer 2024")
```

Chlorophyll A Averages
```{r}
chlA_14 <- SLAP_2024_CHL_A.csv%>%
  filter(MULakeNumber == "014"& endDepth == 0)

chlA_14_avg <- chlA_14 %>%
  group_by(Date)%>%
  summarise(Average_chlA = mean(parameterValue))  #units in ug/L
```

Chlorophyll A Plot
```{r}
chlA_14_avg %>%
  ggplot( aes(x=Date, y=Average_chlA))+
  geom_hline(yintercept = 22.14, size = 2, color = "darkolivegreen")+
  geom_line(aes(group=1), color = "gray76", size = 1.5)+
  geom_point(shape=21, color="black", fill="darkolivegreen", size=8) +
    theme_bw()+
    theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())+
    xlab('Date')+
    ylab('CHL A Concentration (µg/L)')+
    labs(title = "Average Chlorophyll A Concentrations for Summer 2024")
```

## YSI Profiles\

## Temp vs Depth 
```{r}
#after find lake 14, make new column with Fahrenheit instead of Celsius

ysi_14 <- MO_YSIProfileData_2024.csv%>%
  filter(MULakeNumber == "014")

ysi_14te <- ysi_14%>%
  mutate(tempF = (9/5)*temp_degC+32)
```

Temperature vs Depth
```{r}
#treat temp as a function of depth and then reverse later
ggplot(ysi_14te,aes(x=depth_m,y=tempF, group=date))+
  #set up the asthetics, the line must be broken because this is technically discrete data
  geom_line(aes(color =date),size=1.5, linetype=1)+
  #geom_point(aes(color=date))+
  #reverse depth so it starts at zero
  scale_x_reverse()+
  #put the y axis labes on the opposite side so when its flipped it will appear at top
  scale_y_continuous(position="right")+
  #this is how you reverse the look and order or the coordinates for the graph
  coord_flip()+
  theme_bw()+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())+
  xlab('Depth (m)')+
  ylab('Temperature (°F)')+
  theme(legend.position = "inside", legend.position.inside =  c(.90, .20)) #left to right #up and down

```

DO vs Depth
```{r}
#treat temp as a function of depth and then reverse later
ggplot(ysi_14te,aes(x=depth_m,y=doConcentration_mgpL, group=date))+
  #set up the asthetics, the line must be broken because this is technically discrete data
  geom_line(aes(color =date),size=1.5, linetype=1)+
  #geom_point(aes(color=date))+
  #reverse depth so it starts at zero
  scale_x_reverse()+
  #put the y axis labes on the opposite side so when its flipped it will appear at top
  scale_y_continuous(position="right")+
  #this is how you reverse the look and order or the coordinates for the graph
  coord_flip()+
  theme_bw()+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())+
  xlab('Depth (m)')+
  ylab('Dissolved Oxygen Concentration (mg/L)')+
  theme(legend.position = "inside", legend.position.inside =  c(.70, .20)) #left to right #up and down
#300 by 450
```

## Historical Data
First, merge the 2024 data into one data frame with all the parameters
```{r}
folder_2 = "C:/Users/mvhfg/OneDrive - University of Missouri/North_Limnology/priv_lake_rep/Private_Lake_Reports/twf_data"
files_2 = list.files(
  path = folder_2,
  pattern = ".*csv$",
  ignore.case = F,
  full.names = T #provides path to folder
)

# Create a function to read the csv files and return column names
get_column_names <- function(file_path) {
  read_csv(file_path) %>% names()
}

# Get column names from all the files
column_names <- map(files_2, get_column_names)

# Find the common columns in all the files
common_columns <- reduce(column_names, intersect)

# Read and bind data from all the files, keeping only the common columns
combined_24_data <- map_df(files_2, ~read_csv(., col_types = cols(.default = "c"))) %>% select(all_of(common_columns))

```
^ this was practice for gathering historical data and only keeping the matching columns

Cleaning 24 Data to Only show lake 14 and parameters: TN, TP, CHLA COR, SECCHI
```{r}
clean_24yr <- within(combined_24_data, endDepth[parameterType == 'SECCHI'] <- 0)

num_clean_24yr <-clean_24yr %>%
  mutate(endDepth = as.numeric(endDepth))

clean_4_24yr <- num_clean_24yr%>%
  filter(parameterType == 'TN'|parameterType == 'TP'|parameterType == 'CHL_A_COR'|parameterType == 'SECCHI')%>%
  filter(endDepth == 0)%>%
  rename("endDepth_m" = endDepth)

#this is all 2024 data with only the 4 parameters for the graph
```

Put the clean file in the hist_data folder
```{r}
write.csv(clean_4_24yr, "C:/Users/mvhfg/OneDrive - University of Missouri/North_Limnology/priv_lake_rep/Private_Lake_Reports/hist_data/MissouriReservoirs_WaterChemistry_2024.csv",row.names=FALSE)
```

5 years worth of data
```{r}
folder_hist = "C:/Users/mvhfg/OneDrive - University of Missouri/North_Limnology/priv_lake_rep/Private_Lake_Reports/hist_data"
files_hist = list.files(
  path = folder_hist,
  pattern = ".*csv$",
  ignore.case = F,
  full.names = T #provides path to folder
)

# Create a function to read the csv files and return column names
get_column_names <- function(file_path) {
  read_csv(file_path) %>% names()
}

# Get column names from all the files
column_names <- map(files_hist, get_column_names)

# Find the common columns in all the files
common_columns <- reduce(column_names, intersect)

# Read and bind data from all the files, keeping only the common columns
combined_20to24_data <- map_df(files_hist, ~read_csv(., col_types = cols(.default = "c"))) %>% select(all_of(common_columns))
```

Filter to show the private lakes only
```{r}
priv_20to24 <- combined_20to24_data%>%
  filter(MULakeNumber == '014'|MULakeNumber == '021'|MULakeNumber == '096'|MULakeNumber == '018'|
           MULakeNumber == '112'|MULakeNumber == '120'|MULakeNumber == '169'|MULakeNumber == '411'|
           MULakeNumber == '011'|MULakeNumber == '134'|MULakeNumber == '274')%>%
  mutate(endDepth_m = as.numeric(endDepth_m))%>%
  filter(endDepth_m == 0)
```

Graphing Historical Data
```{r}
#Take the annual average and SE of that average
stderr <- function(x) sd(x)/sqrt(length(x))

PrivHistoricalLakes_merged_annual<-priv_20to24%>%
  ungroup()%>%
  mutate(year=year(Date))%>%
  mutate(parameterValue = as.numeric(parameterValue))%>%
  group_by(year,MULakeNumber,waterBody,parameterType,unit)%>%
  dplyr::summarise(parameterValue_mean=mean(parameterValue,na.rm=TRUE),
              parameterValue_sderr=stderr(parameterValue),
              .groups="keep")

#Find the sen slopes for these variables: CHL_A_COR, SECCHI, TN, TP
#group by MULakeNumber, parameterType, unit
PrivHistoricalLakes_merged_annual%>%
  filter(parameterType == 'TN'|parameterType == 'TP'|parameterType == 'CHL_A_COR'|parameterType == 'SECCHI')%>% #pull out these 4 variables of interest
  ungroup()%>%
  group_by(MULakeNumber,parameterType,unit)%>%
  summarize(sensSlope_slope=MTCC.sensSlope(year,parameterValue_mean)$coefficients["Year"],
            sensSlope_intercept=MTCC.sensSlope(year,parameterValue_mean)$coefficients["Intercept"],
            sensSlope_pval=MTCC.sensSlope(year,parameterValue_mean)$pval)%>%
  mutate(significance=ifelse(sensSlope_pval<0.05,"*",""))%>%print(n=Inf)


#for each lake - do this set of 4 figures: TP, TN, CHl, Secchi

#can do it on a loop or just run each one here
lake.id<-"014" #specify the lake
```

Individual Historical Plots for Lake 14
```{r}
#TN specific graphs####
labels_TN<-tibble(x=rep(-Inf,4),y.line=c(350,550,1200,NA),y.label=c((350/2),450,(550+1200)/2,1250),
                  y.label.text=c("0","M","E","HE"))
y.max_TN<-max(1255,max(PrivHistoricalLakes_merged_annual%>%filter(MULakeNumber==lake.id & parameterType=="TN")%>%
                         mutate(upper=parameterValue_mean+parameterValue_sderr)%>%
                         dplyr::select(upper)%>%ungroup()%>%pull(),na.rm=TRUE),na.rm=TRUE)
(gg.TN.lake<-ggplot(data=PrivHistoricalLakes_merged_annual%>%filter(MULakeNumber==lake.id&parameterType=="TN"),
                    aes(x=year,y=parameterValue_mean))+
    geom_hline(yintercept=labels_TN$y.line,linetype=2,color="grey")+
    geom_text(data=labels_TN,aes(x=x,y=y.label,label=y.label.text),hjust=-0.2,color="grey")+
    geom_errorbar(aes(ymin = parameterValue_mean-parameterValue_sderr, ymax = parameterValue_mean+parameterValue_sderr))+
    geom_point(size=2,shape=21,fill="light grey")+
    theme_bw()+
    xlab("Year")+
    ylab(bquote(TN~(mu*g*'/'*L)))+
    scale_y_continuous(limits=c(0,y.max_TN)))

#TP specific graphs####
labels_TP<-tibble(x=rep(-Inf,4),y.line=c(10,25,100,NA),y.label=c((10/2),(10+25)/2,(25+100)/2,120),
                  y.label.text=c("0","M","E","HE"))
y.max_TP<-max(125,max(PrivHistoricalLakes_merged_annual%>%filter(MULakeNumber==lake.id&parameterType=="TP")%>%
                        mutate(upper=parameterValue_mean+parameterValue_sderr)%>%
                        dplyr::select(upper)%>%ungroup()%>%pull(),na.rm=TRUE),na.rm=TRUE)
(gg.TP.lake<-ggplot(data=PrivHistoricalLakes_merged_annual%>%filter(MULakeNumber==lake.id&parameterType=="TP"),
                    aes(x=year,y=parameterValue_mean))+
    geom_hline(yintercept=labels_TP$y.line,linetype=2,color="grey")+
    geom_text(data=labels_TP,aes(x=x,y=y.label,label=y.label.text),hjust=-0.2,color="grey")+
    geom_errorbar(aes(ymin = parameterValue_mean-parameterValue_sderr, ymax = parameterValue_mean+parameterValue_sderr))+
    geom_point(size=2,shape=21,fill="light grey")+
    theme_bw()+
    xlab("Year")+
    ylab(bquote(TP~(mu*g*'/'*L)))+
    scale_y_continuous(limits=c(0,y.max_TP)))

#CHL_A_COR specific graphs####
labels_CHL_A_COR<-tibble(x=rep(-Inf,4),y.line=c(3,9,40,NA),y.label=c((3/2),(3+9)/2,(9+40)/2,43),
                         y.label.text=c("0","M","E","HE"))
y.max_CHL_A_COR<-max(45,max(PrivHistoricalLakes_merged_annual%>%filter(MULakeNumber==lake.id&parameterType=="CHL_A_COR")%>%
                              mutate(upper=parameterValue_mean+parameterValue_sderr)%>%
                              dplyr::select(upper)%>%ungroup()%>%pull(),na.rm=TRUE),na.rm=TRUE)
(gg.CHL_A_COR.lake<-ggplot(data=PrivHistoricalLakes_merged_annual%>%filter(MULakeNumber==lake.id&parameterType=="CHL_A_COR"),aes(x=year,y=parameterValue_mean))+
    geom_hline(yintercept=labels_CHL_A_COR$y.line,linetype=2,color="grey")+
    geom_text(data=labels_CHL_A_COR,aes(x=x,y=y.label,label=y.label.text),hjust=-0.2,color="grey")+
    geom_errorbar(aes(ymin = parameterValue_mean-parameterValue_sderr, ymax = parameterValue_mean+parameterValue_sderr))+
    geom_point(size=2,shape=21,fill="light grey")+
    theme_bw()+

    xlab("Year")+
    ylab(bquote(Chl~italic(a)~(mu*g*'/'*L)))+
    scale_y_continuous(limits=c(0,y.max_CHL_A_COR)))

#SECCHI specific graphs####
labels_SECCHI<-tibble(x=rep(-Inf,4),y.line=c(0.45,1.3,2.6,NA),y.label=c((0.45/2),(0.45+1.3)/2,(1.3+2.6)/2,3),
                      y.label.text=c("HE","E","M","O"))
y.max_SECCHI<-max(3.2,max(PrivHistoricalLakes_merged_annual%>%filter(MULakeNumber==lake.id&parameterType=="SECCHI")%>%
                            mutate(upper=parameterValue_mean+parameterValue_sderr)%>%
                            dplyr::select(upper)%>%ungroup()%>%pull(),na.rm=TRUE),na.rm=TRUE)
(gg.SECCHI.lake<-ggplot(data=PrivHistoricalLakes_merged_annual%>%filter(MULakeNumber==lake.id&parameterType=="SECCHI"),
                        aes(x=year,y=parameterValue_mean))+
    geom_hline(yintercept=labels_SECCHI$y.line,linetype=2,color="grey")+ #Lines for indicating trophic state
    geom_text(data=labels_SECCHI,aes(x=x,y=y.label,label=y.label.text),hjust=-0.2,color="grey")+ #labels for trophic state
    geom_errorbar(aes(ymin = parameterValue_mean-parameterValue_sderr, ymax = parameterValue_mean+parameterValue_sderr))+
    geom_point(size=2,shape=21,fill="light grey")+
    theme_bw()+
    xlab("Year")+
    ylab(bquote(Secchi~(m)))+
    #scale_y_continuous()+
    scale_y_reverse(limits=c(y.max_SECCHI,-0.2)))
```

Making 4 Plots into One Main Figure
```{r}
year.range<-PrivHistoricalLakes_merged_annual%>%filter(MULakeNumber==lake.id)%>%ungroup()%>%
  summarize(min.year=min(year,na.rm=TRUE)-1,max.year=max(year,na.rm=TRUE)+1)%>%slice(1)%>%as.numeric()

IndLakeFigureList<-list(gg.TN.lake+
                      xlab("")+
                      theme(axis.text.x=element_blank(),
                      panel.grid.major = element_blank(), panel.grid.minor = element_blank()
                      )+
                      scale_x_continuous(limits=year.range)+
                      annotate("text",x=Inf,y=Inf,label="A",hjust=1.2,vjust=1.2),
                    gg.TP.lake+
                      xlab("")+
                      #ylab("")+
                      theme(axis.text.x=element_blank(),
                            #axis.text.y=element_blank(),
                            #axis.ticks.x=element_blank(),
                            panel.grid.major = element_blank(), panel.grid.minor = element_blank()
                            )+
                      scale_x_continuous(limits=year.range)+
                      annotate("text",x=Inf,y=Inf,label="B",hjust=1.2,vjust=1.2),
                    gg.CHL_A_COR.lake+
                      #xlab("")+
                      #ylab("")+
                      theme( #axis.text.x=element_blank(),
                        #axis.text.y=element_blank(),
                        #axis.ticks.x=element_blank(),
                        panel.grid.major = element_blank(), panel.grid.minor = element_blank()
                          )+
                      scale_x_continuous(limits=year.range)+
                      annotate("text",x=Inf,y=Inf,label="C",hjust=1.2,vjust=1.2),
                    gg.SECCHI.lake+
                      #xlab("")+
                      #ylab("")+
                      theme(#axis.text.x=element_blank(),
                        #axis.text.y=element_blank(),
                        #axis.ticks.x=element_blank(),
                        panel.grid.major = element_blank(), panel.grid.minor = element_blank()
                        )+
                      scale_x_continuous(limits=year.range)+
                      annotate("text",x=Inf,y=-Inf,label="D",hjust=1.2,vjust=1.2)
)
```

Putting plots on a 2*2 matrix
```{r}
if (!require(patchwork)) {install.packages("patchwork")}
library(patchwork)
(gg.fig3.lake<-patchwork::wrap_plots(IndLakeFigureList,ncol=2,nrow=2)&theme(plot.margin = unit(c(3,3,3,3),"pt")))
```

